Dominance Regions for Imaginary $\bfg$-vectors

### 2020-10-03 dominance_regions.tex
Random minor edits through the first 3 pages.

  \section{Introduction}
  The investigations of this paper are inspired by the dominance order for the tropical labellings for bases of cluster algebras.
  This dominance order is used by Qin to understand how these bases may be deformed in the case when there is a green-to-red sequence of mutations.

@@

  To record sequences of these matrix mutations, we introduce the labelled $n$-regular rooted tree $\TT_n$ with root vertex $t_0$ and associate skew-symmetrizable matrices $B^t$ for $t\in\TT_n$ satisfying:
  \begin{itemize}
    \item $B^{t_0}=B$;
    \item if $t,t'\in\TT_n$ are joined by an edge labelled $k$, then $B^{t'}=\mu_k B^t$.
  \end{itemize}

  (DR: I changed to $\RR$ because it didn't seem like we really cared about integer points at the moment, when we try to link up with minors this will need to be added.)
  For $t\in\TT_n$, define $\phi^t_k:\RR^n\to\RR^n$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ is given by \eqref{eq:mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labelled $k$.
  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labelled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance polytope} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are actually polytopes.
  \end{lemma}

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_i$ for $i\in\ZZ$, with $t_{2j}$ joined to $t_{2j+1}$ by an edge labelled $1$ and $t_{2j-1}$ joined to $t_{2j}$ by an edge labelled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC_{t_i}(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

  Observe that $\phi_{t_1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward mutation 1}
    \phi_{t_1}(\lambda)=\begin{cases} (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1\ge0$;}\\ (-\lambda_1,\lambda_2) & \text{if $\lambda_1<0$;} \end{cases}
  \end{equation}
  and $\phi_{t_2}=\phi_{t_{-2}}^{-1}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} ((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  while $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} (-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

  By an eigenvector of a piecewise-linear map $\phi$, we mean a vector $\lambda$ such that for each $k\ge0$, there exists a scalar $\nu_k$ so that $\phi^k(\lambda)=\nu_k\lambda$.
  \begin{lemma}
    Any nonzero eigenvector of $\phi_{t_2}$ is a positive multiple of one of the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  \end{lemma}
  \begin{proof}
    First observe that the equation $\phi_{t_2}^k(\lambda)=\nu_k\lambda$ cannot be satisfied for all $k\ge0$ unless $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$.
    In this region, $\phi_{t_2}$ is linear with eigenvalue $\nu$ satisfying $\nu^2-(bc-2)\nu+1=0$, i.e. $\nu=\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}$.
    We thus require that 
    \[\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_1=(bc-1)\lambda_1+b\lambda_2 \qquad\text{and}\qquad \frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_2= -c\lambda_1-\lambda_2,\]
    or
    \[\frac{-bc\pm\sqrt{bc(bc-4)}}{2b}\lambda_1=\lambda_2 \qquad\text{and}\qquad \frac{-bc\mp\sqrt{bc(bc-4)}}{2c}\lambda_2=\lambda_1.\]
    As these represent the same relation, the result immediately follows by inspection.
  \end{proof}

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$.
  Also, write $\cI' \subset \RR^2$ for the \emph{dual imaginary cone} (positively) spanned by the vectors $(-bc\pm\sqrt{bc(bc-4)},2c)$.
  \begin{lemma}
    The vectors $(2b, -bc+\sqrt{bc(bc-4)})$ (resp. $(2b, -bc-\sqrt{bc(bc-4)})$) and $(2c, bc+\sqrt{bc(bc-4)})$ (resp. $(2c, bc-\sqrt{bc(bc-4)})$) are perpendicular.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $(-bc+\sqrt{bc(bc-4)})(bc+\sqrt{bc(bc-4)})=-4bc$.
  \end{proof}
  \begin{lemma}
    \label{le:imaginary stability}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have $\phi_{t_{2j}}(\lambda)\in\cI$ and $\phi_{t_{2j+1}}(\lambda)\in\cI'$.
  \end{lemma}
  \begin{proof}
    Note that $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ so the result follows from the cases $j=1$ and $j=0$ respectively.
    The claim is immediate for $\phi_{t_2}$ since the vectors $(2b,-bc\pm\sqrt{bc(bc-4)})$ are eigenvectors.
    Now, we have $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ and so $\phi_{t_1}\big( (2b,-bc\pm\sqrt{bc(bc-4)}) \big)=(-2b,bc\pm\sqrt{bc(bc-4)})$, which is the $\frac{bc\pm\sqrt{bc(bc-4)}}{2c}$-multiple of the boundary vector for $\cI'$ and the result follows for arbitrary $\lambda\in\cI$ by linearity.
  \end{proof}

@@

  The recursion can be written more succinctly as $u_{i+1,\varepsilon}=u_{2,\varepsilon}u_{i,-\varepsilon}-u_{i-1,\varepsilon}$ with the additional initial condition $u_{2,\varepsilon}=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,\ell\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \begin{equation}
      \label{eq:long Chebyshev recursion}
      u_{i+\ell,\varepsilon}=\begin{cases} u_{\ell+1,\varepsilon}u_{i,-\varepsilon}-u_{\ell,-\varepsilon}u_{i-1,\varepsilon} & \text{\upshape if $\ell$ is odd;}\\ u_{\ell+1,-\varepsilon}u_{i,\varepsilon}-u_{\ell,\varepsilon}u_{i-1,-\varepsilon} & \text{\upshape if $\ell$ is even.} \end{cases}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We work by induction on $\ell$, the cases $\ell=0,1$ being tautological and reproducing the defining recursion respectively.
    Using the claim for $j>0$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+\ell+1,\varepsilon}&=\begin{cases} u_{\ell+1,\varepsilon}u_{i+1,-\varepsilon}-u_{\ell,-\varepsilon}u_{i,\varepsilon} & \text{if $\ell$ is odd;}\\ u_{\ell+1,-\varepsilon}u_{i+1,\varepsilon}-u_{\ell,\varepsilon}u_{i,-\varepsilon} & \text{if $\ell$ is even;} \end{cases}\\
      &=\begin{cases} (u_{2,-\varepsilon}u_{\ell+1,\varepsilon}-u_{\ell,-\varepsilon})u_{i,\varepsilon}-u_{\ell+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $\ell$ is odd;}\\ (u_{2,\varepsilon}u_{\ell+1,-\varepsilon}-u_{\ell,\varepsilon})u_{i,-\varepsilon}-u_{\ell+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $\ell$ is even;} \end{cases}\\
      &=\begin{cases} u_{\ell+2,-\varepsilon}u_{i,\varepsilon}-u_{\ell+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $\ell+1$ is even;}\\ u_{\ell+2,\varepsilon}u_{i,-\varepsilon}-u_{\ell+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $\ell+1$ is odd;} \end{cases}
    \end{align*}
    which is the claimed recursion for $\ell+1$.
    These calculations can be reversed to show the result for $\ell<0$.
  \end{proof}

  \begin{lemma}
    \label{le:imaginary transformations}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,-}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,+}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$.
    For $\lambda\in\cI$, the action of $\phi_{t_2}$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_{t_2}(\lambda)=(u_{3,-}\lambda_1+u_{2,+}\lambda_2,-u_{2,-}\lambda_1-u_{1,+}\lambda_2),\]
    and so
    \begin{align*}
      \phi_{t_{2j+2}(\lambda)}&=\phi_{t_2}\phi_{t_{2j}}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_1+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_2,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_1-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_2 \big)\\
      &=(u_{2j+3,+}\lambda_1+u_{2j+2,+}\lambda_2, -u_{2j+2,-}\lambda_1-u_{2j+1,-}\lambda_2),
    \end{align*}
    where the last equality uses \eqref{eq:long Chebyshev recursion} with $i=2j+1$ and $\ell=2$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A=(a_{ij})$ be an invertible $2\times 2$ matrix.
    Under the action of $A$, the region inside $\RR^2$ defined by the inequality $rx+sy\le t$ is transformed to the region defined by the inequality \[\frac{ra_{22}-sa_{21}}{a_{11}a_{22}-a_{12}a_{21}}x + \frac{-ra_{12}+sa_{11}}{a_{11}a_{22}-a_{12}a_{21}}y\le t.\]
  \end{lemma}
  \begin{proof}
    Assume $r,s,t\ne0$, the cases when one of these vanishes follows by the continuity of the transformation.
    Observe that the points $\big(\frac{t}{r},0\big)$ and $\big(0,\frac{t}{s}\big)$ lie on the boundary line of the region defined by the inequality $rx+sy\le t$.
    These are transformed under the action of $A$ into the points $\big(a_{11}\frac{t}{r},a_{21}\frac{t}{r}\big)$ and $\big(a_{12}\frac{t}{s},a_{22}\frac{t}{s}\big)$, respectively, which one readily checks lie along the purported boundary of the image region.
    By the linearity of this transformation of $\RR^2$, we see that either both the domain region and image region contain the origin or both do not, in particular this implies that the inequalities go in the same direction.
  \end{proof}

>>>>

  \section{Introduction}
  The investigations of this paper are inspired by the dominance order on the tropical labelings for bases of cluster algebras.
  This dominance order is used by Qin to understand how these bases may be deformed in the case when there is a green-to-red sequence of mutations.
  In particular, these calculations can be carried out for any acyclic cluster algebra.

@@

  To record sequences of these matrix mutations, we introduce the labeled $n$-regular rooted tree $\TT_n$ with root vertex $t_0$ and associate skew-symmetrizable matrices $B^t$ for $t\in\TT_n$ satisfying:
  \begin{itemize}
    \item $B^{t_0}=B$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $B^{t'}=\mu_k B^t$.
  \end{itemize}

  (DR: I changed to $\RR$ because it didn't seem like we really cared about integer points at the moment, when we try to link up with minors this will need to be added.)
  For $t\in\TT_n$, define $\phi^t_k:\RR^n\to\RR^n$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ is given by \eqref{eq:mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labeled $k$.
  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take the 2-regular tree $\TT_2$ to have vertices labeled $t_i$ for $i\in\ZZ$, with $t_{2j}$ joined to $t_{2j+1}$ by an edge labeled $1$ and $t_{2j-1}$ joined to $t_{2j}$ by an edge labeled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC_{t_i}(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

  Observe that $\phi_{t_1}:=\phi_1^{t_0}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward mutation 1}
    \phi_{t_1}(\lambda)=\begin{cases} (-\lambda_1,c\lambda_1+\lambda_2) & \text{if $\lambda_1\ge0$;}\\ (-\lambda_1,\lambda_2) & \text{if $\lambda_1<0$;} \end{cases}
  \end{equation}
  and $\phi_{t_2}:=\phi_2^{t_1}\phi_1^{t_0}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_{t_2}(\lambda)=\begin{cases} \big((bc-1)\lambda_1+b\lambda_2, -c\lambda_1-\lambda_2\big) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$;}\\ (-\lambda_1, -c\lambda_1-\lambda_2) & \text{if $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2<0$;}\\ (-\lambda_1+b\lambda_2, -\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2\ge 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_1<0$ and $\lambda_2<0$;}\end{cases}
  \end{equation}
  while $\phi_{t_2}^{-1}=\phi_{t_{-2}}$ is the piecewise-linear map given by
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{t_2}^{-1}(\lambda)=\begin{cases} \big(-\lambda_1-b\lambda_2, c\lambda_1+(bc-1)\lambda_2\big) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2\le 0$;}\\ (-\lambda_1-b\lambda_2, -\lambda_2) & \text{if $\lambda_2\le 0$ and $\lambda_1+b\lambda_2>0$;}\\ (-\lambda_1, c\lambda_1-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1\le 0$;}\\ (-\lambda_1,-\lambda_2) & \text{if $\lambda_2>0$ and $\lambda_1>0$;}\end{cases}
  \end{equation}

  By an eigenvector of a piecewise-linear map $\phi$, we will mean a vector $\lambda$ such that for each $k\ge0$, there exists a scalar $\nu_k$ so that $\phi^k(\lambda)=\nu_k\lambda$.
  \begin{lemma}
    Any nonzero eigenvector of $\phi_{t_2}$ is a positive multiple of one of the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  \end{lemma}
  \begin{proof}
    First observe that the equation $\phi_{t_2}^k(\lambda)=\nu_k\lambda$ cannot be satisfied for all $k\ge0$ unless $\lambda_1\ge 0$ and $c\lambda_1+\lambda_2\ge 0$.
    In this region, $\phi_{t_2}$ is linear with eigenvalue $\nu$ satisfying $\nu^2-(bc-2)\nu+1=0$, i.e. $\nu=\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}$.
    We thus require that 
    \[\frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_1=(bc-1)\lambda_1+b\lambda_2 \qquad\text{and}\qquad \frac{bc-2\pm\sqrt{bc(bc-4)}}{2}\lambda_2= -c\lambda_1-\lambda_2,\]
    or, equivalently, that
    \[\frac{-bc\pm\sqrt{bc(bc-4)}}{2b}\lambda_1=\lambda_2 \qquad\text{and}\qquad \frac{-bc\mp\sqrt{bc(bc-4)}}{2c}\lambda_2=\lambda_1.\]
    As these represent the same relation, the result immediately follows by inspection.
  \end{proof}

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  Also, write $\cI' \subset \RR^2$ for the \emph{dual imaginary cone} (positively) spanned by the vectors $\big(-bc\pm\sqrt{bc(bc-4)},2c\big)$.
  \begin{lemma}
    The vectors $\big(2b, -bc+\sqrt{bc(bc-4)}\big)$ (resp. $\big(2b, -bc-\sqrt{bc(bc-4)}\big)$) and $\big(2c, bc+\sqrt{bc(bc-4)}\big)$ (resp. $\big(2c, bc-\sqrt{bc(bc-4)}\big)$) are perpendicular.
  \end{lemma}
  \begin{proof}
    This is immediate from the equality $\big(-bc+\sqrt{bc(bc-4)}\big)\big(bc+\sqrt{bc(bc-4)}\big)=-4bc$.
  \end{proof}
  \begin{lemma}
    \label{le:imaginary stability}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have $\phi_{t_{2j}}(\lambda)\in\cI$ and $\phi_{t_{2j+1}}(\lambda)\in\cI'$.
  \end{lemma}
  \begin{proof}
    Note that $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ so the result follows from the cases $j=1$ and $j=0$, respectively.
    The claim is immediate for $\phi_{t_2}$ since the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$ are eigenvectors and $\phi_{t_2}$ is linear in $\cI$.
    Now, since $b>0$, we have $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ and so $\phi_{t_1}\Big( \big(2b,-bc\pm\sqrt{bc(bc-4)}\big) \Big)=\big(-2b,bc\pm\sqrt{bc(bc-4)}\big)$, which is the $\frac{bc\pm\sqrt{bc(bc-4)}}{2c}$-multiple of the boundary vector for $\cI'$ and the result follows for arbitrary $\lambda\in\cI$ by linearity of $\phi_{t_1}$ in this region.
  \end{proof}

@@

  The recursion can be written more succinctly as $u_{i+1,\varepsilon}=u_{2,\varepsilon}u_{i,-\varepsilon}-u_{i-1,\varepsilon}$ with the additional initial condition $u_{2,\varepsilon}=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,\ell\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \begin{equation}
      \label{eq:long Chebyshev recursion}
      u_{i+\ell,\varepsilon}=\begin{cases} u_{\ell+1,\varepsilon}u_{i,-\varepsilon}-u_{\ell,-\varepsilon}u_{i-1,\varepsilon} & \text{\upshape if $\ell$ is odd;}\\ u_{\ell+1,-\varepsilon}u_{i,\varepsilon}-u_{\ell,\varepsilon}u_{i-1,-\varepsilon} & \text{\upshape if $\ell$ is even.} \end{cases}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We work by induction on $\ell$, the cases $\ell=0,1$ being tautological and reproducing the defining recursion, respectively.
    Using the claim for $j>0$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+\ell+1,\varepsilon}&=\begin{cases} u_{\ell+1,\varepsilon}u_{i+1,-\varepsilon}-u_{\ell,-\varepsilon}u_{i,\varepsilon} & \text{if $\ell$ is odd;}\\ u_{\ell+1,-\varepsilon}u_{i+1,\varepsilon}-u_{\ell,\varepsilon}u_{i,-\varepsilon} & \text{if $\ell$ is even;} \end{cases}\\
      &=\begin{cases} (u_{2,-\varepsilon}u_{\ell+1,\varepsilon}-u_{\ell,-\varepsilon})u_{i,\varepsilon}-u_{\ell+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $\ell$ is odd;}\\ (u_{2,\varepsilon}u_{\ell+1,-\varepsilon}-u_{\ell,\varepsilon})u_{i,-\varepsilon}-u_{\ell+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $\ell$ is even;} \end{cases}\\
      &=\begin{cases} u_{\ell+2,-\varepsilon}u_{i,\varepsilon}-u_{\ell+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $\ell+1$ is even;}\\ u_{\ell+2,\varepsilon}u_{i,-\varepsilon}-u_{\ell+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $\ell+1$ is odd;} \end{cases}
    \end{align*}
    which is the claimed recursion for $\ell+1$.
    These calculations can be reversed to show the result for $\ell<0$.
  \end{proof}

  The following gives the desired expressions for $\phi_{t_i}(\lambda)$ when $\lambda\in\cI$.
  \begin{lemma}
    \label{le:imaginary transformations}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,-}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,+}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$.
    For $\lambda\in\cI$, the action of $\phi_{t_2}$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_{t_2}(\lambda)=(u_{3,-}\lambda_1+u_{2,+}\lambda_2,-u_{2,-}\lambda_1-u_{1,+}\lambda_2),\]
    and so
    \begin{align*}
      \phi_{t_{2j+2}(\lambda)}&=\phi_{t_2}\phi_{t_{2j}}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_1+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_2,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_1-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_2 \big)\\
      &=(u_{2j+3,+}\lambda_1+u_{2j+2,+}\lambda_2, -u_{2j+2,-}\lambda_1-u_{2j+1,-}\lambda_2),
    \end{align*}
    as desired, where the last equality uses \eqref{eq:long Chebyshev recursion} with $i=2j+1$ and $\ell=2$.

@@

  Next we aim to understand how inequalities transform under the action of a linear map, this will be the key tool for describing the dominance regions $\cP(\lambda)$ for $\lambda\in\cI$.
  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A=(a_{ij})$ be an invertible $2\times 2$ matrix.
    Under the action of $A$, the region inside $\RR^2$ defined by the inequality $rx+sy\le t$ is transformed to the region defined by the inequality \[\frac{ra_{22}-sa_{21}}{a_{11}a_{22}-a_{12}a_{21}}x + \frac{-ra_{12}+sa_{11}}{a_{11}a_{22}-a_{12}a_{21}}y\le t.\]
  \end{lemma}
  \begin{proof}
    Assume $r,s,t\ne0$, the cases when one of these vanishes follows by the continuity of the transformation.
    Observe that the points $\big(\frac{t}{r},0\big)$ and $\big(0,\frac{t}{s}\big)$ lie on the boundary line of the region defined by the inequality $rx+sy\le t$.
    These are transformed under the action of $A$ into the points $\big(a_{11}\frac{t}{r},a_{21}\frac{t}{r}\big)$ and $\big(a_{12}\frac{t}{s},a_{22}\frac{t}{s}\big)$, respectively, which one readily checks lie along the purported boundary of the image region.
    By the linearity of the transformation of $\RR^2$ given by the matrix $A$, we see that either both the domain region and image region contain the origin or both do not, in particular this implies that the inequalities go in the same direction.
  \end{proof}
